---
title: "UNODC Reporting"
author: "Connor Flynn"
date: "11/19/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(janitor)
library(ggcorrplot)
```

The main goal of this project is to assess the staus of Small Island Developing States (SIDS) on Criminal Justice-Oriented Sustainable Development Goals (SDG's). Throughout our data-collection process, we encountered significant amounts of mising data. This framed our initial research question to be: Are SIDS reporting data on Criminal Justice-Oriented Sustainable Development Goals less than Non-SIDS. This can inspire insight into the culture and systems of SIDS playing an important role into the survey response rates.

After creating the dataset, we created 3 data visualizations demonstrating survey completion by region for SIDS and Non-SIDS. 



Andrew: Hey Andrew, please comment above each code chunk to say what the code is doing. I've done the first 2 chunks for you


Load in the datasets created from https://dataunodc.un.org/content/ms-data-collection-log by Andrew Nishitomi. The fist is survey response counts for SIDS and the second is survey response counts for Non-SIDS

```{r}
sids_reporting <- read_csv(here("data/UN_CJ_Data_Instruments_SIDS_Reporting.csv"))

non_sids_reporting <- read_csv(here("data/UN_CJ_Data_Instruments _Non_SIDS_Reporting.csv"))

```


Clean the names of the columns of the two dataframes to make them all lower case and seperated by underscores. Using the janitor package, clean_names function
```{r}
sids_reporting <- sids_reporting %>% 
  clean_names()

non_sids_reporting <- non_sids_reporting %>% 
  clean_names()

```


Filter out the years reporting 2017 and 2018 to keep the data frames consistent for the amount of years used.
```{r}
sids_reporting <- sids_reporting %>% 
  filter(year!= 2017 & year!= 2018)

sids_reporting <- sids_reporting %>% 
  filter(country_territory!= "Wallis & Futuna" )

```

Change column name from received to completion status for both sids_reporting and non_sids_reporting
```{r}

colnames(sids_reporting)[colnames(sids_reporting) == "received"] <- "completion_status"

colnames(non_sids_reporting)[colnames(non_sids_reporting) == "received"] <- "completion_status"
```


Converting numeric value to charcter
```{r}

sids_reporting$completion_status <- as.character(sids_reporting$completion_status)

non_sids_reporting$completion_status <- as.character(non_sids_reporting$completion_status)

```

Making characters as 0= "not_completed" and 1="completed"
```{r}

sids_reporting["completion_status"][sids_reporting["completion_status"] == "0"] <- "not_completed"

sids_reporting["completion_status"][sids_reporting["completion_status"] == "1"] <- "completed"


non_sids_reporting["completion_status"][non_sids_reporting["completion_status"] == "0"] <- "not_completed"

non_sids_reporting["completion_status"][non_sids_reporting["completion_status"] == "1"] <- "completed"
```


Counting number of sids that completed and not completed UN Surveys
```{r}
sids_reporting_grouped <- sids_reporting %>% 
  group_by(region, completion_status, data_collection) %>% 
  count()
```

Create a bar graph to display number of surveys completed from each region. As xaxis is Number of surveys completed and yaxis is survey type
```{r}
reporting_status_sids <- ggplot(data = sids_reporting_grouped, aes(x= data_collection, y=n, fill=completion_status)) +
  geom_bar(stat = "identity", position = "dodge") +
  facet_wrap(~region) +
  scale_fill_brewer(palette="Dark2") +
  labs(title = "Reporting Status for 5 UN Surveys from 2019 - 2022, SIDS",
       x = "Survey Type",
       y = "Number of Surveys Completed / Not Completed") +
  theme(axis.text.x = element_text(size = 7, vjust = 1, hjust = 1),
        axis.text.y = element_text(size = 7),
        axis.title.x = element_text(size = 12),
        axis.title.y = element_text(size = 12)) +
  coord_flip()

reporting_status_sids

ggsave(reporting_status_sids, file= here("visualizations/reporting_status_sids.png"), scale=2)



```

Counting number of non-sids that completed and not completed UN Surveys
```{r}
non_sids_reporting_grouped <- non_sids_reporting %>% 
  group_by(region, completion_status, data_collection) %>% 
  count()
```

Show bar graph to show all 5 regions that completed and not completed UN Surveys
```{r}
ggplot(data = non_sids_reporting_grouped, aes(x= data_collection, y=n, fill=completion_status)) +
  geom_bar(stat = "identity", position = "dodge") +
  facet_wrap(~region) +
  scale_fill_brewer(palette="Dark2") +
  labs(title = "Reporting Status for 5 UN Surveys from 2019 - 2022, Non-SIDS",
       x = "Survey Type",
       y = "Number of Surveys Completed / Not Completed") +
  theme(axis.text.x = element_text(size = 7, vjust = 1, hjust = 1),
        axis.text.y = element_text(size = 7),
        axis.title.x = element_text(size = 12),
        axis.title.y = element_text(size = 12))+
  coord_flip()
  
```
Rearranging columns so that data set shows country_territory, data_collection_year, completion_status
```{r}
sids_reporting_matrix <- sids_reporting %>% 
  select(year, completion_status, data_collection, country_territory)

sids_reporting_matrix$data_collection_year <- paste(sids_reporting_matrix$data_collection, "_", sids_reporting_matrix$year)

sids_reporting_matrix <- sids_reporting_matrix %>% 
  select(country_territory, data_collection_year, completion_status)
```




Create Heat Graph that shows completion status per year for every data collection 
```{r}
ggplot(sids_reporting_matrix, aes(x = data_collection_year, y = country_territory, fill = completion_status)) +
  geom_tile(color = "white",
            lwd = 1,
            linetype = 1) +
  coord_fixed() +
  labs(x = "Survey Type & Year",
       y = "SIDS") +
  theme(axis.text.x = element_text(angle = 45, size = 6, vjust = 1, hjust = 1),
        axis.text.y = element_text(size = 9),
        axis.title.x = element_text(size = 12),
        axis.title.y = element_text(size = 12),
        legend.title = element_text(size = 6),
  legend.text = element_text(size = 12)) +
  scale_fill_brewer(palette="Dark2") 
```

